---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# SC2API

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The goal of SC2API is to build a simple and easy-to-use API wrapper for the Blizzard Starcraft II API in the R programming language.

## Installation
You can install the development version of SC2API from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("SamMorrissette/sc2api")
```
## Example Usage

This is a basic example to obtain the MMR (match-maker rating) of the current top 10 players in the North America Grandmaster leaderboard.

First, you must set your own token in the R environment by supplying your own client ID and client secret. For more information on how to obtain these, visit [Getting Started](https://https://develop.battle.net/documentation/guides/getting-started).

```{r Load Library}
library(SC2API)
```

```{r eval=FALSE}
set_token(get_token("YOUR CLIENT ID", "YOUR CLIENT SECRET"))
```

```{r}
data <- get_gm_leaderboard(1, "us") # An argument of "1" corresponds to the NA ladder.
top10 <- data$ladderTeams[1:10] # Extract the top 10 players
playerMMR <- sapply(top10, function(x) x$mmr) # Extract the "mmr" vector from each player. 
print(playerMMR)
```


